$p$-adic precision

### 2014-02-07 padicprec.tex
Differentials of evaluation and interpolation

@@

Beyond sums and products (which can be treated as before), a basic 
operation involving polynomials is Euclidean division. Let us compute 
the differential of it. We start with two polynomials $A$ and $B$ with 
$B \neq 0$. The Euclidean division of $A$ by $B$ is written $A = BQ + R$ 
with $\deg R < \deg B$. Differentiating the above equality we find: 
$dA - dB \cdot Q = B \cdot dQ + dR$,
which implies that $dQ$ and $dR$ are respectively obtained as the 
quotient and the remainder of the Euclidean division of $dA - dB \cdot 
Q$ by $B$. This gives the differential. We note that the discussion 
above extends readily to convergent series (see also 
\cite{Caruso-Lubicz}).

>>>>

\subsubsection*{Evaluation and interpolation}

Beyond sums and products (which can be treated as before), two basic 
operations involving polynomials are evaluation and interpolation.
Evaluation is modelised by the function $(P,x) \mapsto P(x)$ where
$P$ is some polynomial and $x$ some element in $K$. Differentiating
it can be done by computing
\begin{equation}
\label{eq:diffeval}
(P + dP)(x + dx) = P(x + dx) + dP(x + dx) = P(x) + P'(x) dx + dP(x)
\end{equation}
by neglecting terms of order $2$. Here $P'$ denotes the derivative of 
$P$. The differential at $P$ is then the linear map $(dP, dx) \mapsto 
P'(x) dx + dP(x)$.

As for interpolation, we consider a positive integer $d$ and the partial 
function $f : K^{2d} \to K_{< d}[X]$ which maps the tuple $(x_1, y_1, 
\ldots, x_d, y_d)$ to the polynomial $P$ of degree less than $d$ such 
that $P(x_i) = y_i$ for all $i$. The polynomial $P$ exists and is unique 
as soon as the $x_i$'s are pairwise distinct; the above function $f$ is 
then defined on this open set. Furthermore, if $(x_1, y_1, \ldots, x_d, 
y_d)$ is a point in it and $P$ denotes the corresponding interpolation 
polynomial, Eq.~\eqref{eq:diffeval} shows that $d y_i = P'(x_i) dx_i + 
dP(x_i)$ for all $i$. For this, we can compute $dP(x_i)$ from $d x_i$ 
and $d y_i$ and finally recover $dP$ by performing a new interpolation.

@@

Let $A$ and $B$ be two polynomials $B \neq 0$. The Euclidean division of 
$A$ by $B$ is written $A = BQ + R$ with $\deg R < \deg B$. 
Differentiating the above equality we find:
$dA - dB \cdot Q = B \cdot dQ + dR$,
which implies that $dQ$ and $dR$ are respectively obtained as the 
quotient and the remainder of the Euclidean division of $dA - dB \cdot 
Q$ by $B$. This gives the differential. We note that the discussion 
above extends readily to convergent series (see also 
\cite{Caruso-Lubicz}).
